---
title: "Covid91 vaccine study Final2024"
author: "Claudia Cortez!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```


# Introduction
A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

### Males
$H_0$ The COVID vaccine has no effect in males   
$H_A$ The COVID vaccine is effective in males.  

## Methods 
Treatment is a catergorical veriable   
Infected is a catergorial valriable.  
Cat~Cat is the file nI used to receive data.  

## Descriptive Results


### Graphical Descriptive 

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected, data=males)
barchartGC(~infected +treatment, data=males,type="percent")

```
  Based on the graph, it is showing us that the group of males who  took the vaccine are less likely to contract COVID than the ones who received the placebo. 

### Numerical Descriptive
```{r}
table1 <- xtabs(~treatment +infected, data=males)
rowPerc(table1)
colPerc(table1)
```

  This table is showing that 74% of the males who experienced the placebo got Covid19. While only 26% of those who got the vaccines contacted the virus. 

### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for the chi-square test is 0.000000000000002, and Fisher's exact test provides a p-value of 0.0000000000000002 So we reject the null hypothesis. 

The odds ratio of 28.1:1 indicates that people who took the placebo are 28.1 times more likely to contract COVID-19 compared to thoes who got the vaccine. 

### Conclusion
 The graph demonstrates that men who were vaccinated had a significantly lower risk of contracting COVID -19 compared to those who received the placebo. The data reveals that 74% of males who took the placebo contracted Covid-19, whereas only 26% of males who received the vaccine contacted the virus. The chi-square test shows a p-value 0.0000000000000002, and Fisher's exact test has the p value of 0.000000000000000002,both of which are well below 0.05 , indicating significant results. the odds ratio of 2.8:1 suggests that those who took the placebo are 28.1 times more likely to contract COVID-19 compared to those who received the vaccine.  
 
 
 

# Females
$H_0$ The COVID vaccine has no effect in females  
$H_A$ The COVID vaccine is effective in females  

## Methods 
Treatment is a categorical veriable  
Sex ( females )is a categorical variable  
Cat~Cat is the file nI used to receive data. 

## Descriptive Results


### Graphical Descriptive 

```{r}
females=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected, data=females)
barchartGC(~infected +treatment, data=females,type="percent")

```
  
  
  
  This shows that females that received the vaccine are less likely to get Covid than people who got the placebo. 



## Results


```{r}
femaleData = filter (FinalData,sex == "F")
```


## Descriptive Results

### Graphical Descriptive 

```{r}
barchartGC(~infected+treatment,data=femaleData)
barchartGC(~infected+treatment,data=femaleData, type="percent")
```


  This graph is showing that from the all female group who contracted COVID, the ones who received the vaccine had less symptoms than those with the placebo while the all the female group that did not contract COVID showed no difference between those who were vaccinated and not vaccinated. 
### Numerical Descriptive
```{r}
table2 <- xtabs(~infected + treatment, data=femaleData)
rowPerc(table2)
colPerc(table2)
```

  The data from this table shows that from the female group that took the placebo were infected 65% with COVID 19, females that got the vaccine had 35% got COVID. 
  
  
### Inferential Results
```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

  Based on this chisq test, the p-value is 0.00000002, showing that there is a significant difference, concluding that we are rejecting the null hypothesis. 
For the fisher exact test the p-value is 0.00000001. This is lower than 0.05 so we reject the null. 

The odd ratio is a 1.85:1 this shows that females that took the placebo are 1.86 less likely to contract COVID-19 compared to those that were vaccinated.

###LGBTQ
$H_0$ The COVID vaccine has no effect in LGBTQ  
$H_A$ The COVID vaccine is effective in LGBTQ  

## Methods 
Treatment is a categorical veriable  
Sex ( LGBTQ )is a categorical variable    
Cat~Cat is the file used to receive data  

## Descriptive Results

### Graphical Descriptive


```{r}
LGBTQData = filter (FinalData,LGBTQ == "gay")
```


```{r}
barchartGC(~infected+treatment,data=LGBTQData)
barchartGC(~infected+treatment,data=LGBTQData, type="percent")
```

  This graph is showing that within the LGBTQ subgroup who were vaccinated have a low risk of getting COVID -19  compared to those who got the placebo. 
  
  
### Numerical Descriptive

```{r}
table3 <- xtabs(~ treatment  + infected, data=LGBTQData)
rowPerc(table3)
colPerc(table3)
```

  This table is showing that within the LGBTQ placebo group they experienced 80% of symptoms while the the vaccinated group experienced 20% of symptoms. 

### Inferential Results
```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

  The chisq test shows that the p-value is 0.000002. Which is lower 0.05 concluding that we are rejecting the null hypothesis.  

In the fisher exact test, p-value is 0.00000007 also lower than 0.05
showing a difference between the two groups, which is another example of rejecting the null hypothesis in this group.


# Druggies 

$H_0$ The COVID vaccine has no effect in Drug User      
$H_A$ The COVID vaccine is effective in Drug User    

## Methods 
Treatment is a categorical veriable    
Sex (Drug User)is a categorical variable     
Cat~Cat is the file used to receive data   

## Descriptive Results

### Graphical Descriptive


```{r}
DruggiesData = filter(FinalData, DrugUser == "yes")
```


```{r}
barchartGC(~infected+treatment,data=DruggiesData)
barchartGC(~infected+treatment,data=DruggiesData, type="percent")
```

This graph us showing that from the drug user group that took the placebo were less likely to get COVID, compared to the ones who received the vaccine. 

### Numerical Descriptive
```{r}
table4 <- xtabs(~ treatment + infected, data=DruggiesData)
rowPerc(table4)
colPerc(table4)
```
This table is showing that within the placebo drug user group, they experience 8.33% symptoms while the vaccinated group experienced 91.67% of symptoms. 


### Inferential Results
```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```


Based on this chisq test, the p-value is 0.00000002842 

The fisher exact test showed a p-value of 0.00000232, which is showing that there is a  odds ratio of about 0.12%, suggesting that the odds of this occuring are lower in one group compared to the other, the 95% confidence ineravl supports this 

# Overall Results and Conclusions

In conclusion, the effectiveness of the COVID vaccine varies between subgroups. Staring with the all male group, we are rejecting the null hypothesis. Based on the bar chart, numerical graph, and the chisq and fisher tests, we can conclude that men who received the COVID vaccine and contracted it ultimately experience less symptoms than those who received the placebo. The men that did not get vaccinated and did not contract COVID showed little to no difference between eachother. In the group that did receive the placebo had an infection rate of 2.55%, and those of them who got the vaccine had an infection rate of 1.15%, showing that the vaccine is effective in men, therefore we are rejecting the null hypothesis for the male group. 

Moving onto the all female group, we are also rejecting the null hypothesis. Based on all of the tests we have run, we can conclude that women who received the COVID vaccine and contracted it also experienced less symptoms than the women who received the placebo. The female group that did not get vaccinated and did not contract COVID also showed little to no difference compared to each other. Within the group that did receive the placebo, they had an infection rate of 2.04%. The group that received the vaccine had an infection rate of 1.37%, concluding that the vaccine is effective in women, therefore rejecting the null hypothesis for the female subgroup.

Comparing the male and female subgroups, women had a smaller percentage of infection with the COVID vaccine and placebo than then men by 0.22% and 0.51%. It is not a big difference, but so far within the data analysis women experience less symptoms with the COVID vaccine. 

In the LGBTQ subgroup, we are rejecting the null hypothesis. Based on our tests, we can conclude that the COVID vaccine is effective in this subgroup, but negatively. We found out that the LGBTQ people who received the COVID vaccine and contracted it experienced significant symptoms than those who received the placebo. The LGBTQ group that did not get vaccinated and did not contract COVID were almost exactly the same, showing no difference. In the LGBTQ group that did receive the placebo had an infection rate of 0.66% and those who did receive the vaccine had an infection rate of 3.54%, a very significant difference between the two, therefore rejecting the null hypothesis since the vaccine had an effect on the group.

Finally, for the drug user group, they received the same result as the LGBTQ subgroup. We are rejecting the null hypothesis due to the drug users experiencing significantly more symptoms than the placebo group, showing that the COVID vaccine did have an effect, but not a positive one In the group of drug users that were vaccinated and contracted COVID, they had a 3.06% infection rate. The placebo group had a 0.4% infection rate showing a very big difference between the two and how much symptoms each group got, once again, we are rejecting the null.

Comparing all of the tests from most positive COVID affects to the more negative COVID affects, the female had the best results with the COVID vaccine, while the LGBTQ subgroup had it the worst, experiencing the most symptoms while having COVID even with the vaccine. Males had almost the same results as the females, just a little less and drug users had fewer symptoms by only 0.48%. I would recommend that only males and females that are not drug users or LGBTQ to get the COVID vaccine since the effect was more positive than it was for the other two subgroups. 


