---
title: "Covid91 vaccine study Final2024"
author: "Claudia Cortez!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
$H_0$: The Covid921 vaccine is not effetive in males    
$H_A$  The Covid921 vavccine is not effective in females 

### Methods 

Treatment is cat   
sex is cat 
helper file we are using is CAT- CAT 


### Description Results 


### Graphical Results 
```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males) 
barchartGC(~infected + treatment,data=males,type="percent")
```
 
 This shows that males who got the vaccine are less likely to get covid than people who the placcebo. 

### Numerical 
```{r}
table1 <- xtabs(~treatment + infected, data=males)
rowPerc(table1)
colPerc(table1)
```
  This shows that 74% of males who took the palcebo got covid19, as were 26% males got covid19


### Inferential Results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```






### Conclusion 



## Females

## Males
$H_0$: The Covid921 vaccine is not effective in males    
$H_A$  The Covid921 vavccine is not effective in females 




## LGBTQ
$H_0$: The Covid921 vaccine is not effetive in males    
$H_A$  The Covid921 vavccine is not effective in females 



## Druggies
$H_0$: The Covid921 vaccine is not effetive in drug users.    
$H_A$  The Covid921 vavccine is not effective in drug users.  


# Overall Results and Conclusions