---
title: "Covid91 vaccine study Final2024"
author: "Claudia Cortez!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```


# Introduction
A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

### Males
$H_0$ The COVID vaccine has no effect in males   
$H_A$ The COVID vaccine is effective in males.  

## Methods 
Treatment is a catergorical veriable   
Infected is a catergorial veriable.  
Cat~Cat is the file I used to capture the data.  

## Descriptive Results


### Graphical Descriptive 

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected, data=males)
barchartGC(~infected +treatment, data=males,type="percent")

```

  Based on the graph, it is showing us that the group of males who took the vaccine are less likely to contract COVID than the ones who received the placebo. 

### Numerical Descriptive
```{r}
table1 <- xtabs(~treatment +infected, data=males)
rowPerc(table1)
colPerc(table1)
```

  This table is showing that 74% of the males who experienced the placebo got Covid19. While only 26% of those who got the vaccines contacted the virus. 

### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for the chi-square test is 0.000000000000002, and Fisher's exact test provides a p-value of 0.0000000000000002 So we reject the null hypothesis. 

The odds ratio of 28.1:1 indicates that people who took the placebo are 28.1 times more likely to contract COVID-19 compared to thoes who got the vaccine. 

### Conclusion
 The graph demonstrates that men who were vaccinated had a significantly lower risk of contracting COVID -19 compared to those who received the placebo. The data reveals that 74% of males who took the placebo contracted Covid-19, whereas only 26% of males who received the vaccine contacted the virus. The chi-square test shows a p-value 0.0000000000000002, and Fisher's exact test has the p value of 0.000000000000000002,both of which are well below 0.05 , indicating significant results. the odds ratio of 2.8:1 suggests that those who took the placebo are 28.1 times more likely to contract COVID-19 compared to those who received the vaccine.  
 
 

# Females
$H_0$ The COVID vaccine has no effect in females  
$H_A$ The COVID vaccine is effective in females  

## Methods 
Treatment is a categorical veriable  
Sex ( females )is a categorical variable  
Cat~Cat is the file nI used to receive data. 

## Descriptive Results


### Graphical Descriptive 
```{r}
femaleData = filter (FinalData,sex == "F")
```

```{r}
females=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected, data=females)
barchartGC(~infected +treatment, data=females,type="percent")

```
  
  
  
  This shows that females that received the vaccine are less likely to get Covid19 than people who got the placebo. 

### Numerical Descriptive
```{r}
table2 <- xtabs(~infected + treatment, data=femaleData)
rowPerc(table2)
colPerc(table2)
```

  The data from this table shows that from the female group that took the placebo were infected 65% with COVID 19, females that got the vaccine had 35% got COVID. 
  
  
### Inferential Results
```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

  Based on this chisq test, the p-value is 0.00000002, showing that there is a significant difference, concluding that we are rejecting the null hypothesis. 
For the fisher exact test the p-value is 0.00000001. This is lower than 0.05 so we reject the null. 

The odd ratio is a 1.85:1 this shows that females that took the placebo are 1.86 less likely to contract COVID19 compared to those that were vaccinated.

### Conclusion
 The findings indicate that females who received the vaccine have a lower risk of contracting Covid19 compared to those who received the placebo. According to the data presented in the table, 65% of females in the placebo group were infected with Covid19, whereas among those who received the vaccine, 35% were infected. The chi-square test yielded a p-value of 0.00000002, indicating a significant difference and leading to rejection of the null hypothesis.

Similarly, the Fisher exact test resulted in a p-value of 0.00000001, which is below the typical significance level of 0.05, also supporting rejection of the null hypothesis.

The odds ratio of 1.85:1 suggests that females in the placebo group are 1.85 times more likely to contract Covid19 compared to those who were vaccinated.

### LGBTQ  

$H_0$ The COVID vaccine has no effect in LGBTQ  
$H_A$ The COVID vaccine is effective in LGBTQ    

## Methods 
Treatment is a categorical veriable  
Sex ( LGBTQ )is a categorical variable    
Cat~Cat is the file used to receive data  

## Descriptive Results

### Graphical Descriptive


```{r}
LGBTQData = filter (FinalData,LGBTQ == "gay")
```


```{r}
barchartGC(~infected+treatment,data=LGBTQData)
barchartGC(~infected+treatment,data=LGBTQData, type="percent")
```

  This graph is showing that within the LGBTQ subgroup who were vaccinated have a low risk of getting COVID -19  compared to those who got the placebo. 
  
  
### Numerical Descriptive

```{r}
table3 <- xtabs(~ treatment  + infected, data=LGBTQData)
rowPerc(table3)
colPerc(table3)
```

  This table is showing that within the LGBTQ placebo group they experienced 80% of symptoms while the the vaccinated group experienced 20% of symptoms. 

### Inferential Results
```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

  The chisq test shows that the p-value is 0.000002. Which is lower 0.05 concluding that we are rejecting the null hypothesis.  

In the fisher exact test, p-value is 0.00000007 also lower than 0.05
showing a difference between the two groups, which is another example of rejecting the null hypothesis in this group.

### Conclusion
The graph indicates that LGBTQ individuals who received the vaccine have a significantly lower risk of contracting Covid19 compared to those who received the placebo. According to the table, 80% of the LGBTQ placebo group experienced symptoms, whereas only 20% of the vaccinated group experienced symptoms. The chi-square test resulted in a p-value of 0.000002, which is less than 0.05, leading to rejection of the null hypothesis.

Similarly, the Fisher exact test yielded a p-value of 0.00000007, also below the significance level of 0.05, indicating a notable difference between the two groups and supporting rejection of the null hypothesis within this subgroup.

# Druggies 

$H_0$ The COVID vaccine has no effect in Drug User      
$H_A$ The COVID vaccine is effective in Drug User    

## Methods 
Treatment is a categorical veriable    
Sex (Drug User)is a categorical variable     
Cat~Cat is the file used to receive data   

## Descriptive Results

### Graphical Descriptive


```{r}
DruggiesData = filter(FinalData, DrugUser == "yes")
```


```{r}
barchartGC(~infected+treatment,data=DruggiesData)
barchartGC(~infected+treatment,data=DruggiesData, type="percent")
```

  This graph us showing that from the drug user group that took the placebo were less likely to get COVID, compared to the ones who received the vaccine. 

### Numerical Descriptive
```{r}
table4 <- xtabs(~ treatment + infected, data=DruggiesData)
rowPerc(table4)
colPerc(table4)
```

  This table is showing that within the placebo drug user group, they experience 8.33% symptoms while the vaccinated group experienced 91.67% of symptoms. 


### Inferential Results
```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The chisq test shows that the p-value is 0.0000000001. Which is lower 0.05 concluding that we are rejecting the null hypothesis.  

In the fisher exact test, p-value is 0.000000000003 also lower than 0.05
showing a difference between the two groups, which is another example of rejecting the null hypothesis in this group.

The odd ratio 0.087:1, the drug user that used the palcebo is 0.08 less likely then the one who took the vaccine. 

### Conclusion 
This graph illustrates that among drug users, those who received the placebo were more likely to contract COVID-19 compared to those who received the vaccine. According to the table, within the placebo group of drug users, 8.33% experienced symptoms, while 91.67% of the vaccinated group experienced symptoms.

The chi-square test yielded a p-value of 0.0000000001, indicating a significant difference and leading to the rejection of the null hypothesis (p < 0.05). Similarly, the Fisher exact test resulted in a p-value of 0.000000000003, further supporting the rejection of the null hypothesis, demonstrating a clear distinction between the two groups within this context.

The odds ratio of 0.087:1 indicates that drug users who received the placebo were 0.087 times less likely to contract COVID-19 compared to those who received the vaccine. At the end of all , the vaccine is effective to Drugusers however they would get covid tenn those who took the palcebo. 
